Quantitative review of the effects of triazole and benzimidazole fungicides on Fusarium head blight and wheat yield in Brazil

### 2017-10-24 article.Rmd
test web
audpc2

Franklin Jackson Machado<sup>1</sup>, Flávio Martins Santana<sup>2</sup>, Douglas Lau<sup>2</sup> and Emerson Medeiros Del Ponte<sup>1</sup>

<sup>1</sup> Departamento de Fitopatologia, Universidade Federal de Viçosa, , Viçosa 36570-000, MG, Brazil

<sup>2</sup> Embrapa Trigo, Passo Fundo 70770-901 RS, Brazil;

@@

> Triazole and benzimidazole fungicides have been used for controlling Fusarium head blight (FHB) in wheat for over two decades. In Brazil, it was only during the last five years that uniform fungicide trials for FHB control have been established yearly, thus contributing to a new large body of fungicide efficacy data for this country. A systematic review of both peer- and non-peer reviewed studies on chemical control conducted since 2000 in Brazil was performed. The fungicide treatments of interest were the triazoles tebuconazole (TEBU<sub>1x</sub> and TEBU<sub>2x</sub>) and propiconazole (PROP<sub>2x</sub>), and the benzimidazole carbendazim (CARB2x). These four treatments were applied most twice (2x), the first at early to mid-flowering and the second 7 to 10 days later.  Only TEBU was tested as one (1x) or two applications. For these treatments, there were 35 trials reporting FHB index and 48 reporting mean yield, for which separate meta-analyses were performed. Network meta-analytic models were fitted to the data of the log of the means of FHB index for each fungicide and for the non-treated check. The meta-analytic estimates were used to obtain control efficacy (C̅), or percent disease reduction relative to the non-treated check. The absolute mean difference (D̅) in yield (kg/ha) between the fungicide-treated and the non-treated check plots was also estimated. Yield response relative to the non-treated check (Y̅) was also estimated based on the difference in the logs of the means of yield between fungicide-treated and non-treated check. The TEBU<sub>1x</sub>, TEBU<sub>2x</sub> and CARB2x treatments performed similarly with regards control efficacy (C̅ = 59%, 53% and 55%, respectively), and although better than PROP<sub>2</sub> (47%), the difference was marginally significant. Mean yield response was highest for TEBU2x, (D̅ = 558 kg/ha ,Y̅ = 19.2%) followed by PROP<sub>2x</sub> (497 kg/ha, 16.0%), TEBU<sub>1x</sub> (457 kg/ha, 17.3%), and CARB<sub>2x</sub> (456 kg/ha, 12.8%). For an average 2016 scenario of fungicide plus application costs (F<sub>c</sub> = $18 U.S./ha) and average wheat price (Wp = $215 U.S./MT), the probability of breaking even on the financial investment in the four treatments ranged from 59% to 63%. For additional 140 scenarios (four fungicides) created based on the combination of five WP ($133 to 266 U.S./MT) and seven F<sub>c</sub> ($5 to 35 U.S./ha), the probability of breaking even was >50% for all but two scenarios. Results can be useful as a guide for planning future trials and provides a baseline and first step towards optimizing FHB management in Brazil.

Keywords: *Fusarium graminearum*; *Triticum aestivum*; chemical control; multi-treatment meta-analysis.

@@

Fusarium head blight (FHB) caused mainly by the *Fusarium graminearum* species complex (FGSC) is among the most important fungal diseases of wheat worldwide (Goswami and Kistler 2004; McMullen et al. 2012). In the presence of airborne inoculum, the first infections may occur during the onset of flowering if wet conditions prevail. The earlier the infection, the greater the damage in grain yield because infected kernels show reduced size, light weight and a shriveled appearance (McMullen et al. 2012). However, post-flowering infection, although less damaging to grain yield, may contribute to the accumulation of mycotoxins in the grain (Del Ponte et al. 2007). FGSC members are able to produce a wide range of mycotoxins, but those of greatest concern are the trichothecenes, especially deoxynivalenol (DON), which is commonly found in commercial grain at unacceptable levels for both human and animal consumption (Del Ponte et al. 2012; McMullen et al. 2012). Maximum tolerated levels of DON in small grain and foods have been established in many countries (Van Egmond et al. 2007) including Brazil (ANVISA 2011).

@@

Three response variables of interest were obtained: 1) percent control efficacy (C̅) of FHB index relative to the non-treated check. There were four studies that reported only FHB incidence. For those, FHB index was estimated from incidence using a complementary log-log model developed with data collected in the same region across 160 fields (Spolti et al. 2015); 2) absolute difference (D̅) in wheat yield between the fungicide treatment and the non-treated check. The latter metric can be used when the variation in the reference treatment (e.g. non-treated check) across studies is not large (Madden and Paul, 2011; Paul et al. 2010). In our study, the calculated coefficients of variation were 131% and 38% for FHB index and yield in the non-treated check, respectively, thus supporting the use of D for yield data; 3) The percent yield increase (Y̅) relative to the non-treated check was also calculated (Paul et al. 2010). The estimated variances (V) (within-study or sampling variance) of the log or non-transformed means of treatment were estimated as described (Paul et al. 2010, 2008).

@@

where Yi is the vector of *L* (log of the means of severity or yield) or mean yield for the four treatments plus the non-treated check for the ith study, µ is a vector representing mean of Yi across all studies, Σ is a 5 × 5 between-study variance-covariance matrix, and Si is within-study variance-covariance matrix for the ith study. N(•) indicates a multivariate normal distribution. The elements of Si were incorporated into the model fitting procedure as weights calculated as the inverse function of the within-study variance for each treatment from each study as described (Paul et al. 2008). An unstructured Σ matrix was used and the models were fitted to the data with a maximum-likelihood parameter. The model was fitted using the ‘metafor’ package (Viechtbauer, 2010) of R (R Core Team 2012). To test for inconsistency of the network, for both the estimates of the log of the means of FHB index and non-transformed means of yield, all trials were categorized according to their design or the common set of treatments tested in the trial. In total, eight different designs were found for testing the fungicide effects on both variables. A significant (*P* = 0.05) interaction of the treatment and design factors, tested in the meta-analytic model as described, evaluated based on the Wald test statistic, was an indication of inconsistency (Piepho et al. 2015; Madden et al. 2016).

@@

Since the log of the response ratios equal the difference of logs, the differences between pairs of the elements in µ, as well as their standard errors and 95% confidence intervals were obtained (Paul et al. 2008). Thus, the overall mean log of the response ratio (L̅) for each fungicide treatment relative to the non-treated check was estimated as L ̅" = " "μ"  ̂_"Treat "  "- " "μ"  ̂_"Check" , where "μ"  ̂_"Treat"   and "μ"  ̂_"Check"  are estimated mean log FHB index or yield for a treatment (Treat) and the non-treated check (Check), respectively. Overall mean percent FHB control and mean yield increase and their confidence intervals were obtained by back-transforming "L"  ̅_IND and "L"  ̅_YLD and the upper and lower limits of the confidence intervals around "L"  ̅_"IND"  and "L"  ̅_"YLD"  as "C"  ̅ = (1-(exp("L"  ̅_"IND"  ))×100) and Y ̅ = ((exp("L"  ̅_"YLD" )-1)×100), respectively. The mean difference in yield (D̅) was estimated directly by the difference between the estimates for the fungicide treatment and the non-treated check (Madden et al., 2016).

@@

*Probability of breaking-even fungicide application cost.* The mean yield difference (D̅) and the between-study variance (τ2) estimated by the meta-analytic models were used to calculate the probability of breaking even when considering the fungicide application cost (Fc = fungicide plus application costs). This was calculated as the cumulative standard-normal distribution function of [(D̅ – FC/WP)/ √(τ ̂^2 )] as in previous studies (Paul et al. 2011; Salgado et al. 2014; Tedford et al. 2017). The operational costs were fixed at $10 U.S./ha. Average prices of triazoles (PROP or TEBU) and CARB were $8.9 and $7.9 U.S./ha, respectively, considering the average exchange rate of $3.16 BRL = $1 USD during September 2016. The average wheat price (WP) used in our analysis was $215 U.S./MT ($680 BRL) according to estimates for that same growing season. Given the variability of prices and costs at the farm-level and across years, we further calculated the probabilities of breaking even for a range of WP ($133 to 266 U.S./MT) and Fc ($5 to 35 U.S./ha) values, totaling 140 scenarios.

>>>>

```{r, include=FALSE}
Sys.setlocale("LC_ALL", 'pt_BR.UTF-8')
```

Franklin Jackson Machado, Flávio Martins Santana, Douglas Lau and Emerson Medeiros Del Ponte

1 Departamento de Fitopatologia, Universidade Federal de Viçosa, , Viçosa 36570-000, MG, Brazil

2 Embrapa Trigo, Passo Fundo 70770-901 RS, Brazil;

@@

> Triazole and benzimidazole fungicides have been used for controlling Fusarium head blight (FHB) in wheat for over two decades. In Brazil, it was only during the last five years that uniform fungicide trials for FHB control have been established yearly, thus contributing to a new large body of fungicide efficacy data for this country. A systematic review of both peer- and non-peer reviewed studies on chemical control conducted since 2000 in Brazil was performed. The fungicide treatments of interest were the triazoles tebuconazole (TEBU1x and TEBU2x) and propiconazole (PROP2x), and the benzimidazole carbendazim (CARB2x). These four treatments were applied most twice (2x), the first at early to mid-flowering and the second 7 to 10 days later.  Only TEBU was tested as one (1x) or two applications. For these treatments, there were 35 trials reporting FHB index and 48 reporting mean yield, for which separate meta-analyses were performed. Network meta-analytic models were fitted to the data of the log of the means of FHB index for each fungicide and for the non-treated check. The meta-analytic estimates were used to obtain control efficacy (C̅), or percent disease reduction relative to the non-treated check. The absolute mean difference (D̅) in yield (kg/ha) between the fungicide-treated and the non-treated check plots was also estimated. Yield response relative to the non-treated check (Y̅) was also estimated based on the difference in the logs of the means of yield between fungicide-treated and non-treated check. The TEBU1x, TEBU2x and CARB2x treatments performed similarly with regards control efficacy (C̅ = 59%, 53% and 55%, respectively), and although better than PROP2x (47%), the difference was marginally significant. Mean yield response was highest for TEBU2x, (D̅ = 558 kg/ha ,Y̅ = 19.2%) followed by PROP2x (497 kg/ha, 16.0%), TEBU1x (457 kg/ha, 17.3%), and CARB2x (456 kg/ha, 12.8%). For an average 2016 scenario of fungicide plus application costs (Fc = $18 U.S./ha) and average wheat price (Wp = $215 U.S./MT), the probability of breaking even on the financial investment in the four treatments ranged from 59% to 63%. For additional 140 scenarios (four fungicides) created based on the combination of five WP ($133 to 266 U.S./MT) and seven Fc ($5 to 35 U.S./ha), the probability of breaking even was >50% for all but two scenarios. Results can be useful as a guide for planning future trials and provides a baseline and first step towards optimizing FHB management in Brazil.

Keywords: Fusarium graminearum; Triticum aestivum; chemical control; multi-treatment meta-analysis.

@@

Fusarium head blight (FHB) caused mainly by the Fusarium graminearum species complex (FGSC) is among the most important fungal diseases of wheat worldwide (Goswami and Kistler 2004; McMullen et al. 2012). In the presence of airborne inoculum, the first infections may occur during the onset of flowering if wet conditions prevail. The earlier the infection, the greater the damage in grain yield because infected kernels show reduced size, light weight and a shriveled appearance (McMullen et al. 2012). However, post-flowering infection, although less damaging to grain yield, may contribute to the accumulation of mycotoxins in the grain (Del Ponte et al. 2007). FGSC members are able to produce a wide range of mycotoxins, but those of greatest concern are the trichothecenes, especially deoxynivalenol (DON), which is commonly found in commercial grain at unacceptable levels for both human and animal consumption (Del Ponte et al. 2012; McMullen et al. 2012). Maximum tolerated levels of DON in small grain and foods have been established in many countries (Van Egmond et al. 2007) including Brazil (ANVISA 2011).

@@

Three response variables of interest were obtained: 1) percent control efficacy ("C"  ̅) of FHB index relative to the non-treated check. There were four studies that reported only FHB incidence. For those, FHB index was estimated from incidence using a complementary log-log model developed with data collected in the same region across 160 fields (Spolti et al. 2015); 2) absolute difference (D ̅) in wheat yield between the fungicide treatment and the non-treated check. The latter metric can be used when the variation in the reference treatment (e.g. non-treated check) across studies is not large (Madden and Paul, 2011; Paul et al. 2010). In our study, the calculated coefficients of variation were 131% and 38% for FHB index and yield in the non-treated check, respectively, thus supporting the use of D for yield data; 3) The percent yield increase (Y ̅) relative to the non-treated check was also calculated (Paul et al. 2010). The estimated variances (V) (within-study or sampling variance) of the log or non-transformed means of treatment were estimated as described (Paul et al. 2010, 2008).

@@

where Yi is the vector of "L" _ (log of the means of severity or yield) or mean yield for the four treatments plus the non-treated check for the ith study, µ is a vector representing mean of Yi across all studies, Σ is a 5 × 5 between-study variance-covariance matrix, and Si is within-study variance-covariance matrix for the ith study. N(•) indicates a multivariate normal distribution. The elements of Si were incorporated into the model fitting procedure as weights calculated as the inverse function of the within-study variance for each treatment from each study as described (Paul et al. 2008). An unstructured Σ matrix was used and the models were fitted to the data with a maximum-likelihood parameter. The model was fitted using the ‘metafor’ package (Viechtbauer, 2010) of R (R Core Team 2012). To test for inconsistency of the network, for both the estimates of the log of the means of FHB index and non-transformed means of yield, all trials were categorized according to their design or the common set of treatments tested in the trial. In total, eight different designs were found for testing the fungicide effects on both variables. A significant (P = 0.05) interaction of the treatment and design factors, tested in the meta-analytic model as described, evaluated based on the Wald test statistic, was an indication of inconsistency (Piepho et al. 2015; Madden et al. 2016).

@@

Since the log of the response ratios equal the difference of logs, the differences between pairs of the elements in µ, as well as their standard errors and 95% confidence intervals were obtained (Paul et al. 2008). Thus, the overall mean log of the response ratio (L ̅) for each fungicide treatment relative to the non-treated check was estimated as L ̅" = " "μ"  ̂_"Treat "  "- " "μ"  ̂_"Check" , where "μ"  ̂_"Treat"   and "μ"  ̂_"Check"  are estimated mean log FHB index or yield for a treatment (Treat) and the non-treated check (Check), respectively. Overall mean percent FHB control and mean yield increase and their confidence intervals were obtained by back-transforming "L"  ̅_IND and "L"  ̅_YLD and the upper and lower limits of the confidence intervals around "L"  ̅_"IND"  and "L"  ̅_"YLD"  as "C"  ̅ = (1-(exp("L"  ̅_"IND"  ))×100) and Y ̅ = ((exp("L"  ̅_"YLD" )-1)×100), respectively. The mean difference in yield (D̅) was estimated directly by the difference between the estimates for the fungicide treatment and the non-treated check (Madden et al., 2016).

@@

*Probability of breaking-even fungicide application cost.* The mean yield difference (D ̅) and the between-study variance (τ ̂^2) estimated by the meta-analytic models were used to calculate the probability of breaking even when considering the fungicide application cost (Fc = fungicide plus application costs). This was calculated as the cumulative standard-normal distribution function of [(D̅ – FC/WP)/ √(τ ̂^2 )] as in previous studies (Paul et al. 2011; Salgado et al. 2014; Tedford et al. 2017). The operational costs were fixed at $10 U.S./ha. Average prices of triazoles (PROP or TEBU) and CARB were $8.9 and $7.9 U.S./ha, respectively, considering the average exchange rate of $3.16 BRL = $1 USD during September 2016. The average wheat price (WP) used in our analysis was $215 U.S./MT ($680 BRL) according to estimates for that same growing season. Given the variability of prices and costs at the farm-level and across years, we further calculated the probabilities of breaking even for a range of WP ($133 to 266 U.S./MT) and Fc ($5 to 35 U.S./ha) values, totaling 140 scenarios.
